---
title: "GC Candidate Tree Maps"
author: "Kelsey McGurrin Dexter H. Locke"
date: "`r format(Sys.time())`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  

```{r include=FALSE}
## 0 set up: load libraries, custom functions, set defaults

# load libraries
# packages we'll be using
packs <- c(
  'tidyverse'         # a must have!
  # , 'tidylog'         # makes things very verbose for 2x checking 
  # , 'magrittr'        # all of the pipes
  # , 'janitor'         # cleans things up
  , 'sf'              # spatial support
  #, 'tidycensus'      # Census access
  , 'mapview'         # webmaps
  # , 'tictoc'          # times things
  # , 'beepr'           # makes noises
  )         

# check for all of the libraries
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))  
}

# load them
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)



# keep random things consistent
# set.seed(19870630) # needed?


# # redlining colors
# holc_pal <- c('#92BC6B' # green
#               , '#92C7C9' # blue
#               , '#E7DC6B' # yellow
#               , '#E47D67' # red
#               #, '#A9A9A9'
#               ) # dark gray)

# redlining colors
holc_pal_group <- c('#92BC6B' # green
                    , '#92C7C9' # blue
                    , '#E7DC6B' # yellow
                    , '#E47D67' # red
                    #, '#A9A9A9' # dark gray)
                    )

# holc_pal_f<- c('#92BC6B' # green
#               , '#92C7C9' # blue
#               , '#E7DC6B' # yellow
#               , '#E47D67' # red
#               , '#A9A9A9'
#               , '#000000')


# custom function for "Not In"
`%nin%` <- Negate(`%in%`)


# fixes mapview
mapviewOptions(fgb = FALSE)

```


```{r include=FALSE}
## 1 read in HOLC polygons

#define new groups for HOLC
(holc_lu <- 
  tribble(
    ~grade, ~grade_group,
    "A",    "AB",
    "B",    "AB",
    "C",    "C",
    "D",    "D"
    ,    NA,     "industrial"
  ))


#join with shapefile
(holc <- st_read('output_tables/baci_holc_polygons_2024-02-20.gpkg', as_tibble = TRUE) %>% 
    left_join(holc_lu,by="grade") %>%
    select(area_id,geom,grade,grade_group))

```



```{r include=FALSE}
## 2 read in tree data

trees <-  
  st_read('../grand_challenges_large/candidate_street_trees_2024-06-28.gpkg') 


trees$area_id<- replace_na(trees$area_id,0)
trees$grade<- replace_na(trees$grade,"undesignated")

```


# Summary Table
```{r include=FALSE}

# make table with spp as rows, number of neighborhoods and trees as columns
tab<-trees %>%
  st_drop_geometry() %>%
  group_by(SPP,grade_group) %>%
  summarize(n_trees=n(),n_neighborhoods=length(unique(area_id))) %>%
  drop_na() %>%
  pivot_wider(id_cols=SPP,names_from = grade_group,values_from = c(n_trees,n_neighborhoods)) %>%
  mutate(n_trees_other=sum(n_trees_industrial,n_trees_undesignated,na.rm = T),.keep="unused",.after=n_trees_D) %>%
  select(-c(n_neighborhoods_industrial,n_neighborhoods_undesignated))


```
```{r echo=FALSE}
DT::datatable(tab)
```



# Acer buergerianum (Trident maple)
n = `r nrow(trees|>filter(SPP=="Acer buergerianum"))`
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer buergerianum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Acer platanoides (Norway maple)
n = `r nrow(trees|>filter(SPP=="Acer platanoides"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer platanoides") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Acer rubrum (Red maple)
n = `r nrow(trees|>filter(SPP=="Acer rubrum"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer rubrum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Acer saccharinum (Silver maple)
n = `r nrow(trees|>filter(SPP=="Acer saccharinum"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer saccharinum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Acer saccharum (Sugar maple)
n = `r nrow(trees|>filter(SPP=="Acer saccharum"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer saccharum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Aesculus glabra (Ohio buckeye)
n = `r nrow(trees|>filter(SPP=="Aesculus glabra"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Aesculus glabra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Amelanchier spp. (Serviceberry spp.)
n = `r nrow(trees|>filter(SPP=="Amelanchier spp."))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Amelanchier spp.") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Betula nigra (River birch)
n = `r nrow(trees|>filter(SPP=="Betula nigra"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Betula nigra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Carpinus caroliniana (American hornbeam)
n = `r nrow(trees|>filter(SPP=="Carpinus caroliniana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Carpinus caroliniana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Carya illinoinensis (Pecan)
n = `r nrow(trees|>filter(SPP=="Carya illinoinensis"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Carya illinoinensis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Celtis laevigata (Sugarberry)
n = `r nrow(trees|>filter(SPP=="Celtis laevigata"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Celtis laevigata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Celtis occidentalis (Hackberry)
n = `r nrow(trees|>filter(SPP=="Celtis occidentalis"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Celtis occidentalis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Cercis canadensis (Eastern redbud)
n = `r nrow(trees|>filter(SPP=="Cercis canadensis"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cercis canadensis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Cladrastis kentukea (Yellowwood)
n = `r nrow(trees|>filter(SPP=="Cladrastis kentukea"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cladrastis kentukea") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Cornus florida (Flowering dogwood)
n = `r nrow(trees|>filter(SPP=="Cornus florida"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cornus florida") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Cornus kousa (Kousa dogwood)
n = `r nrow(trees|>filter(SPP=="Cornus kousa"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cornus kousa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Ginkgo biloba (Ginkgo)
n = `r nrow(trees|>filter(SPP=="Ginkgo biloba"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ginkgo biloba") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Gleditsia triacanthos (Honeylocust)
n = `r nrow(trees|>filter(SPP=="Gleditsia triacanthos"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Gleditsia triacanthos") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Ilex opaca (American holly)
n = `r nrow(trees|>filter(SPP=="Ilex opaca"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ilex opaca") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Juniperus virginiana (Eastern redcedar)
n = `r nrow(trees|>filter(SPP=="Juniperus virginiana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Juniperus virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Lagerstroemia indica (Crapemyrtle)
n = `r nrow(trees|>filter(SPP=="Lagerstroemia indica"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Lagerstroemia indica") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Liquidambar styraciflua (Sweetgum)
n = `r nrow(trees|>filter(SPP=="Liquidambar styraciflua"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Liquidambar styraciflua") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Liriodendron tulipifera (Tuliptree)
n = `r nrow(trees|>filter(SPP=="Liriodendron tulipifera"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Liriodendron tulipifera") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

#  Magnolia grandiflora (Southern magnolia)
n = `r nrow(trees|>filter(SPP=="Magnolia grandiflora"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Magnolia grandiflora") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

#  Magnolia virginiana (Sweetbay magnolia)
n = `r nrow(trees|>filter(SPP=="Magnolia virginiana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Magnolia virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

#  Nyssa sylvatica (Blackgum)
n = `r nrow(trees|>filter(SPP=="Nyssa sylvatica"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Nyssa sylvatica") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Pinus virginiana (Virginia pine)
n = `r nrow(trees|>filter(SPP=="Pinus virginiana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Pinus virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Pinus strobus (Eastern white pine)
n = `r nrow(trees|>filter(SPP=="Pinus strobus"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Pinus strobus") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```


# Platanus occidentalis (Sycamore)
n = `r nrow(trees|>filter(SPP=="Platanus occidentalis"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Platanus occidentalis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Platanus x acerifolia (London planetree)
n = `r nrow(trees|>filter(SPP=="Platanus x acerifolia"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Platanus x acerifolia") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Prunus cerasifera (Cherry plum)
n = `r nrow(trees|>filter(SPP=="Prunus cerasifera"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Prunus cerasifera") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Prunus serotina (Black cherry)
n = `r nrow(trees|>filter(SPP=="Prunus serotina"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Prunus serotina") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus alba (White oak)
n = `r nrow(trees|>filter(SPP=="Quercus alba"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus alba") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus bicolor (Swamp white oak)
n = `r nrow(trees|>filter(SPP=="Quercus bicolor"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus bicolor") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus falcata (Southern red oak)
n = `r nrow(trees|>filter(SPP=="Quercus falcata"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus falcata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus lyrata (Overcup oak)
n = `r nrow(trees|>filter(SPP=="Quercus lyrata"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus lyrata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus macrocarpa (Bur oak)
n = `r nrow(trees|>filter(SPP=="Quercus macrocarpa"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus macrocarpa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus michauxii (Swamp chesnut oak)
n = `r nrow(trees|>filter(SPP=="Quercus michauxii"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus michauxii") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus palustris (Pin oak)
n = `r nrow(trees|>filter(SPP=="Quercus palustris"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus palustris") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus phellos (Willow oak)
n = `r nrow(trees|>filter(SPP=="Quercus phellos"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus phellos") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Quercus rubra (Northern red oak)
n = `r nrow(trees|>filter(SPP=="Quercus rubra"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus rubra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Robinia pseudoacacia (Black locust)
n = `r nrow(trees|>filter(SPP=="Robinia pseudoacacia"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Robinia pseudoacacia") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Salix nigra (Black willow)
n = `r nrow(trees|>filter(SPP=="Salix nigra"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Salix nigra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Syringa reticulata (Japanese tree lilac)
n = `r nrow(trees|>filter(SPP=="Syringa reticulata"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Syringa reticulata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Taxodium distichum (Baldcypress)
n = `r nrow(trees|>filter(SPP=="Taxodium distichum"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Taxodium distichum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Thuja occidentalis (Eastern arborvitae)
n = `r nrow(trees|>filter(SPP=="Thuja occidentalis"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Thuja occidentalis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Tilia americana (American linden)
n = `r nrow(trees|>filter(SPP=="Tilia americana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Tilia americana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Tilia tomentosa (Silver linden)
n = `r nrow(trees|>filter(SPP=="Tilia tomentosa"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Tilia tomentosa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Ulmus americana (American elm)
n = `r nrow(trees|>filter(SPP=="Ulmus americana"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ulmus americana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```

# Zelkova serrata (Japanese zelkova)
n = `r nrow(trees|>filter(SPP=="Zelkova serrata"))`
```{r echo=FALSE, warning=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Zelkova serrata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade"
          , col.regions = holc_pal_group
          , layer.name = 'HOLC grade'
          , alpha.regions = .4
          ) 

```