---
title: "05_make_candidate_maps"
author: "Kelsey McGurrin Dexter H. Locke"
date: "`r format(Sys.time())`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  

```{r include=FALSE}
## 0 set up: load libraries, custom functions, set defaults

# load libraries
# packages we'll be using
packs <- c(
  'tidyverse'         # a must have!
  # , 'tidylog'         # makes things very verbose for 2x checking 
  # , 'magrittr'        # all of the pipes
  # , 'janitor'         # cleans things up
  , 'sf'              # spatial support
  #, 'tidycensus'      # Census access
  , 'mapview'         # webmaps
  # , 'tictoc'          # times things
  # , 'beepr'           # makes noises
  )         

# check for all of the libraries
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packs, rownames(installed.packages())))  
}

# load them
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)



# keep random things consistent
# set.seed(19870630) # needed?


# # redlining colors
# holc_pal <- c('#92BC6B' # green
#               , '#92C7C9' # blue
#               , '#E7DC6B' # yellow
#               , '#E47D67' # red
#               #, '#A9A9A9'
#               ) # dark gray)

# redlining colors
holc_pal_group <- c('#92BC6B' # green
                    # , '#92C7C9' # blue
                    , '#E7DC6B' # yellow
                    , '#E47D67' # red
                    # , '#A9A9A9' # dark gray)
                    )

# holc_pal_f<- c('#92BC6B' # green
#               , '#92C7C9' # blue
#               , '#E7DC6B' # yellow
#               , '#E47D67' # red
#               , '#A9A9A9'
#               , '#000000')


# custom function for "Not In"
`%nin%` <- Negate(`%in%`)


# fixes mapview
mapviewOptions(fgb = FALSE)

```


```{r include=FALSE}
## 1 read in HOLC polygons

#define new groups for HOLC
(holc_lu <- 
  tribble(
    ~grade, ~grade_group,
    "A",    "AB",
    "B",    "AB",
    "C",    "C",
    "D",    "D",
    NA,     "industrial"
  ))


#read in shapefile, add new column, drop industrial category
(holc <- st_read('output_tables/baci_holc_polygons_2024-02-20.gpkg', as_tibble = TRUE) %>% 
    left_join(holc_lu,by="grade") %>%
    filter(grade_group %nin% "industrial") %>%
    select(area_id,geom,grade_group))

```



```{r include=FALSE}
## 2 read in tree data

trees <-  
  st_read('../grand_challenges_large/candidate_street_trees_2024-04-11.gpkg')
  

```


# Acer buergerianum (Trident maple)
n = 256
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer buergerianum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Acer platanoides (Norway maple)
n = 166
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer platanoides") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Acer rubrum (Red maple)
n = 2822
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer rubrum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Acer saccharinum (Silver maple)
n = 145
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer saccharinum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Acer saccharum (Sugar maple)
n = 634
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Acer saccharum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Aesculus glabra (Ohio buckeye)
n = 54
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Aesculus glabra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Amelanchier spp. (Serviceberry spp.)
n = 469
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Amelanchier spp.") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Betula nigra (River birch)
n = 63
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Betula nigra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Carpinus caroliniana (American hornbeam)
n = 174
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Carpinus caroliniana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Celtis laevigata (Sugarberry)
n = 43
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Celtis laevigata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Celtis occidentalis (Hackberry)
n = 455
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Celtis occidentalis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Cercis canadensis (Eastern redbud)
n = 1120
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cercis canadensis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Cladrastis kentukea (Yellowwood)
n = 168
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cladrastis kentukea") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Cornus florida (Flowering dogwood)
n = 347
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cornus florida") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Cornus kousa (Kousa dogwood)
n = 84
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Cornus kousa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Ginkgo biloba (Ginkgo)
n = 574
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ginkgo biloba") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Gleditsia triacanthos (Honeylocust)
n = 14
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Gleditsia triacanthos") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Ilex opaca (American holly)
n = 109
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ilex opaca") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Juniperus virginiana (Eastern redcedar)
n = 72
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Juniperus virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Lagerstroemia indica (Crapemyrtle)
n = 768
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Lagerstroemia indica") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Liquidambar styraciflua (Sweetgum)
n = 931
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Liquidambar styraciflua") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Liriodendron tulipifera (Tuliptree)
n = 129
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Liriodendron tulipifera") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

#  Magnolia grandiflora (Southern magnolia)
n = 52
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Magnolia grandiflora") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

#  Magnolia virginiana (Sweetbay magnolia)
n = 41
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Magnolia virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

#  Nyssa sylvatica (Blackgum)
n = 676
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Nyssa sylvatica") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Pinus virginiana (Virginia pine)
n = 13
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Pinus virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Platanus occidentalis (Sycamore)
n = 106
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Pinus virginiana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Platanus x acerifolia (London planetree)
n = 1229
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Platanus x acerifolia") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Prunus cerasifera (Cherry plum)
n = 105
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Prunus cerasifera") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus alba (White oak)
n = 144
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus alba") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus bicolor (Swamp white oak)
n = 667
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus bicolor") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus falcata (Southern red oak)
n = 53
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus falcata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus lyrata (Overcup oak)
n = 62
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus lyrata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus macrocarpa (Bur oak)
n = 50
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus macrocarpa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus michauxii (Swamp chesnut oak)
n = 44
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus michauxii") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus palustris (Pin oak)
n = 364
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus palustris") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus phellos (Willow oak)
n = 1061
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus phellos") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Quercus rubra (Northern red oak)
n = 495
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Quercus rubra") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Robinia pseudoacacia (Black locust)
n = 48
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Robinia pseudoacacia") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Syringa reticulata (Japanese tree lilac)
n = 224
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Syringa reticulata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Taxodium distichum (Baldcypress)
n = 174
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Taxodium distichum") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Thuja occidentalis (Eastern arborvitae)
n = 162
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Thuja occidentalis") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Tilia americana (American linden)
n = 288
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Tilia americana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Tilia tomentosa (Silver linden)
n = 150
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Tilia tomentosa") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Ulmus americana (American elm)
n = 408
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Ulmus americana") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```

# Zelkova serrata (Japanese zelkova)
n = 1356
```{r echo=FALSE}
## 3 make maps

trees |>
  filter(SPP == "Zelkova serrata") |>
  mapview(  zcol = 'mean_temp_C'
          , layer.name = 'mean_temp_C'
          , alpha.regions = .9
          ) +
  mapview( holc, zcol = "grade_group"
          , col.regions = holc_pal_group
          , layer.name = 'grade group'
          , alpha.regions = .4
          ) 

```